---
title: "Homework 3"
author: "Vihar Desu"
output: github_document
---

# Problem 1

```{r imports, echo=FALSE, message=FALSE}
library(tidyverse)
library(p8105.datasets)
data("instacart")
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

The instacart dataset provides useful information about what their users order at grocery stores including how frequently or regularly they use the instacart service to purchase their grocery items. Relevant columns may include "order_dow", "order_hour_of_day", "days_since_prior_order", etc. It also lends itself to analysis of what product categories that these purchased grocery items fall in with columns like "aisle" and "department". This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns of information. We have information for `r instacart %>% distinct(order_id) %>% count()` of orders, purchased accross `r instacart %>% distinct(department) %>% count()` of departments and `r instacart %>% distinct(aisle) %>% count()` aisles from `r instacart %>% distinct(user_id) %>% count()` users.

### Instacart Dataset
```{r, instacart, echo=FALSE, message=FALSE}
knitr::kable(head(instacart),"simple", caption = "Source: Instacart")
```

### Aisles
```{r, aisles, echo=TRUE, message=FALSE}
aisles = instacart %>%
  count(aisle) %>% 
  arrange(desc(n))
```
There are `r instacart %>% distinct(aisle) %>% count()` distinct aisles, with the most popular category ordered was fresh vegetables appearing 150,609 times in orders. Fresh fruits are a close second at 150,473 times in orders.

```{r, iaisles_table, echo=FALSE, message=FALSE}
knitr::kable(head(aisles),"simple", caption = "Source: Instacart")
```



As we can see in the graph above, two most popular categories by far are the fresh fruits and fresh vegetables. The other popular categories include such items as milk, yogurt and sparkling water. In general though, most aisles do trend toward the same amount of ordered items well below 40,000 in total.
```{r, echo=TRUE, message=FALSE}
instacart %>%
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  ggtitle("Number Items Ordered by Aisle") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Baking Ingredients, Dog Food Care and Packaged Vegtables/Fruits
As we can see, the most popular items in each category are light brown sugar, snack sticks chicken, and organic baby spinach. The number of ordered items differs largely across these three categories, indicating that packaged vegetables and fruits are widely more popular than all of the ranked items combined in this table).
```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable("simple", caption = "Instacart Dataset")
```


### Pink Lady Apples and Coffee Ice Cream
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream") ) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  mutate(
    mean_hour = paste0(as.character(floor(mean_hour)),":",as.character(floor((mean_hour - floor(mean_hour)) * 60)))
  ) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  rename(c("Product Name" = "product_name", "Sunday"="0", "Monday"="1", "Tuesday"="2", "Wednesday"="3", "Thursday"="4", "Friday"="5", "Saturday"="6")) %>% 
  knitr::kable()

```




```{r, echo=TRUE, message=FALSE}

library(ggridges)
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

head(weather_df)

```

```{r, echo=TRUE, message=FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

```{r, echo=TRUE, message=FALSE}



ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

