---
title: "Homework 3"
author: "Vihar Desu"
output: github_document
---

# Problem 1

```{r imports, echo=FALSE, message=FALSE}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
data("instacart")
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

The instacart dataset provides useful information about what their users order at grocery stores including how frequently or regularly they use the instacart service to purchase their grocery items. Relevant columns may include "order_dow", "order_hour_of_day", "days_since_prior_order", etc. It also lends itself to analysis of what product categories that these purchased grocery items fall in with columns like "aisle" and "department". This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns of information. We have information for `r instacart %>% distinct(order_id) %>% count()` of orders, purchased accross `r instacart %>% distinct(department) %>% count()` of departments and `r instacart %>% distinct(aisle) %>% count()` aisles from `r instacart %>% distinct(user_id) %>% count()` users.

### Instacart Dataset
```{r, instacart, echo=FALSE, message=FALSE}
knitr::kable(head(instacart),"simple", caption = "Source: Instacart")
```

### Aisles
```{r, aisles, echo=TRUE, message=FALSE}
aisles = instacart %>%
  count(aisle) %>% 
  arrange(desc(n))
```
There are `r instacart %>% distinct(aisle) %>% count()` distinct aisles, with the most popular category ordered was fresh vegetables appearing 150,609 times in orders. Fresh fruits are a close second at 150,473 times in orders.

```{r, iaisles_table, echo=FALSE, message=FALSE}
knitr::kable(head(aisles),"simple", caption = "Source: Instacart")
```

As we can see in the graph above, two most popular categories by far are the fresh fruits and fresh vegetables. The other popular categories include such items as milk, yogurt and sparkling water. In general though, most aisles do trend toward the same amount of ordered items well below 40,000 in total.
```{r, echo=TRUE, message=FALSE}
instacart %>%
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  ggtitle("Number Items Ordered by Aisle") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Baking Ingredients, Dog Food Care and Packaged Vegtables/Fruits
As we can see, the most popular items in each category are light brown sugar, snack sticks chicken, and organic baby spinach. The number of ordered items differs largely across these three categories, indicating that packaged vegetables and fruits are widely more popular than all of the ranked items combined in this table).
```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable("simple", caption = "Instacart Dataset")
```


### Pink Lady Apples and Coffee Ice Cream
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream") ) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  mutate(
    mean_hour = paste0(as.character(floor(mean_hour)),":",as.character(floor((mean_hour - floor(mean_hour)) * 60)))
  ) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  rename(c("Product Name" = "product_name", "Sunday"="0", "Monday"="1", "Tuesday"="2", "Wednesday"="3", "Thursday"="4", "Friday"="5", "Saturday"="6")) %>% 
  knitr::kable()

```

## Problem 2

-pivot_longer activity data
-split day into weekend and weekdays


```{r, accel_data, echo=FALSE, message=TRUE}

accel_data = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_minute",
    names_prefix = "activity_",
    values_to = "activity_level"
  ) %>% 
  mutate(
    weekday = case_when(
      day == "Monday" | day == "Tuesday" | day == "Wednesday" | day == "Thursday" | day == "Friday" ~ TRUE,
      day == "Saturday" | day =="Sunday" ~ FALSE
    )
  ) %>% 
  select(
    day_id,
    week_id = week,
    weekday,
    day,
    minute = activity_minute,
    everything()
  )

```

As mentioned in the homework assignment, this dataset contains five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). This dataset contains activity level data of this male on one minute intervals for all 5 weeks that his experiment was conducted. There are `r nrow(accel_data)` rows of data, which represent entries for "activity_level" temporally categorized by columns "day_id", "week_id", "weekday", "day" and "minute". 


```{r}

accel_data_totals = 
  accel_data %>% 
  group_by(day_id, week_id, day, weekday) %>% 
  summarize(activity_total = sum(activity_level)) %>% 
  arrange(desc(activity_total))

week_plot = 
  accel_data_totals %>% 
  ggplot(aes(x = day, y = activity_total)) + 
  geom_point() +
  geom_boxplot() +
  labs(
    title = "Activity by Day of the Week",
    x = "Day of the Week (n = 5/day)",
    y = "Activity Level",
    caption = "Activity Level Variability"
  )
week_plot

```

From the results of the graph, we can see that there is an observable difference in this individual's activity variability on Saturdays. Monday's tended to be the high activity level day, while on average the individual tended to have an activity level around 4e+05. What's also consistent is that the individual's activity level mirrored a sort of wave like plot, around his average activity level over time. 

```{r}
daily_plot = 
  accel_data_totals %>% 
  ggplot(aes(x = day_id, y = activity_total)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Activity throughout Study Interval",
    x = "Days of Study (n = 35)",
    y = "Activity Level",
    caption = "Activity Level Over Time"
  )


daily_plot
```


-35 lines
-one line one day
-x-axis is 1440 minutes
-y-axis is activity level

```{r}



plot = 
  accel_data %>% 
  group_by(day_id) %>% 
  ggplot(aes(x = minute, y = activity_level, color = day)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "bottom")
plot


```

